---
title: "LAGOS Spatial Analysis"
author: "Elaina Passero"
date: "9/27/2019"
output: html_document
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())

#Load in lagos
#lagos <- lagosne_load()

#Grab the lake centroid info
#lake_centers <- lagos$locus

load('lake_centers.Rdata')
```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r }
#Your code here
IA_IL <- states %>%
  filter(name %in% c("Iowa","Illinois")) %>%
  st_transform(crs=2163)

mapview(IA_IL)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r}
IA_IL_lakes <- spatial_lakes[IA_IL,] # spatial clip

# count lakes
nrow(IA_IL_lakes)
nrow(minnesota_lakes)
```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
# get outlines of states
iowa_minnesota <- states %>%
  filter(name %in% c("Iowa","Minnesota")) %>%
  st_transform(2163)

# spatially join lakes to their states
iowa_minnesota_lakes <- spatial_lakes %>%
  st_join(.,iowa_minnesota) %>%
  filter(!is.na(name))

ggplot(iowa_minnesota_lakes,aes(lake_area_ha))+
  geom_histogram()+
  facet_wrap(~name,ncol=1)+ # name comes from spatial join; ncol=1 stacks them
  scale_x_log10()

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares

```{r}
mapview(IA_IL_lakes,zcol="lake_area_ha")
```


## 5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 

```{r}
mapview(IA_IL_lakes,zcol="lake_perim_meters")
```

Lake perimeter describes the spatial extent of the lakes and reservoirs.
